Phenomenological Robotics

### 2011-12-16 draft.tex
Wrote about Todes vs. Kuipers. Wrote about Spinozism.

% 1.E. Robots have different bodies, different perceptions and different needs
%      from ours.
% (Source: Thesis journal, pages 006 to 015, 14/12/2010).

@@

Gibson~\cite{gibson2000} propose that the process of learning is one of
discrimination rather than association, that is, through the refinement of the
sensorimotor knowledge acquired through the selection of patterns in our
interaction with the world. According to her, this selection process happens
based on two criteria:

@@

>>>>

% 1.E. Robots have different bodies, different perceptions and different needs
%      from ours.
% (Source: Thesis journal, pages 006 to 015, 14/12/2010).
% Note: it might make more sense to propose the importance of the robots' bodies
% after introducing Todes (section 3.A). I'm not sure, however, of how to
% restructure the text to accomplish that.

@@

Gibson and Pick~\cite{gibson2000} propose that the process of learning is one of
discrimination rather than association, that is, through the refinement of the
sensorimotor knowledge acquired through the selection of patterns in our
interaction with the world. According to them, this selection process happens
based on two criteria:

@@

% 3. In Robotics, Phylosophy and Related Work.

\section{Phenomenology and robotics}

% 3.A. Kuipers vs. Todes.
% (Source: \cite{todes2001}, page xix).
% (Source: Thesis journal, pages 42 to 42, 05/01/2011).

Dreyfus, in his introduction to Todes' \emph{Body and World}~\cite{todes2001},
describes the fundamental role played by the body in structuring the
spatiotemporal field that defines human experience

\begin{quotation}
  \dots the structure of the active body produces our unified experience of
  space and time. Since the body moves forward more effectively than backwards,
  it opens a \emph{horizontal field} that organizes experience in what can be
  coped with directly, what can be reached with effort, and what is over the
  perceptual horizon. Furthermore, the front/back asymmetry of the active body
  \dots organizes the temporal field.
\end{quotation}

% Reviewing Dreyfus introduction to Todes' Body and World, I realized that his
% work contributes a lot more than just the importance of the body and its
% orientation. According to Dreyfus, he goes further on improving the concepts
% of intentional arc and maximal grip from Merleau-Ponty. It might be an
% interesting source of ideas, but I'm not sure they would fit in this text.

Pierce and Kuipers~\cite{pierce1997} developed an algorithm to obtain a robot's
perception model and its possible actions in an unknown environment. Despite the
importance given to the robot's body, they make a clear distinction between the
mind (agent) and the body (robot), seeing the first as learning to control the
second.

% Not present in the outline: Spinozism and how it counters Pierce and Kuipers.
% I'm not sure if the way the text below is structured adds much to the point.
% (Source: \cite{lemmen1997}).
% ATTENTION: I had access to this thesis via Porfirio Silva, who downloaded it
% from the author's website in 2002 (no longer available). According to his
% note, the author asks not to be cited without permission. I found no trace of
% the author online, with the exception of a paper from 1996 with the email
% ronaldl@cogs.susx.ac.uk. Chapters 6 and 7 might be worth a full read, as they
% verse on ``(Artificial) Life and Dynamical Theory'' and on ``Body and Soul''
% (``Non-Cartesian Robots'').

Lemmen~\cite{lemmen1997} developed what he called a ``Spinozist'' framework for
understanding cognition. Rooted on the work of Merleau-Ponty, one of its main
points is the proposal of leaving Cartesianism behind, stating that

\begin{quotation}
  For the Spinozist, mind and body form a concrete unity.
\end{quotation}
